Evaluating the Capabilities of the Lattice Boltzmann Method for non-Newtonian and Free-surface Flow towards Applications in Wellbore Cementing

### 2017-03-02 response-to-reviewers.tex

\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}

\begin{document}

\section{Response to reviewers}

Thank you very much for your assistance in reviewing this paper. We are sure that with your help, this has become a much better work. We hope that the attached edited copy of "Numerical Investigation of the Accuracy, Stability, and Efficiency of Lattice Boltzmann Methods in Simulating Non-Newtonian Flow" will answer any questions previously raised, and will be satisfactory for publication.

The following is a list of all comments recieved, and how they have been addressed in the updated version:

\end{document}

### 2017-03-02 src/stability-paper.tex
copied template for cover letter to editor and response to reviewers. fixed grammatical and spelling errors in the text

There are a number of fluids in science and engineering applications that can be classified as non-Newtonian, such as pastes, slurries, molten plastics, polymer solutions, dyes, varnishes, suspensions, and some biomedical liquids such as blood all behave in a non-Newtonian manner~\cite{bohme1987non}.
Of all of the different non-Newtonian behaviors that exist, there are two models under which much of the behaviors may be idealized: yield stress fluids and power-law fluids.
Yield stress fluids, also known as Bingham plastics, do not flow until a threshold value of stress, referred to as its yield stress, is exceeded.
Yield stress flow has been found to be relevant in many applictions, such as flow of pastes, paints, muds, molten plastics and metals, and in some cases, blood~\cite{wang2011lattice}.

@@

Due to its simplicity and computational efficiency, the most common collision operator is the Bhatnagar--Gross--Krook (BGK) operator.
BGK consists of a single relaxation time and is a linear relaxation of particle distributions toward equilibrium.
The BGK collison operator for the $i^{th}$ discrete velocity is expressed as:

@@

Despite the utility of the BGK collision operator, it does have a few drawbacks.
For example, in low viscosity fluids the BGK operator results in an overrelaxation of particle distributions toward quasiequilibrium.
It is well known that when large nonequilibrium distributions exist in the LBM approximation that overrelaxation can result in nonphysical oscillations that are slow to decay~\cite{brownlee2007stability,dellar2003incompressible}.
To illustrate this, consider a flow in which there is a sharp spatial gradient in either $\rho$ or $\mvel$ at $\pos$.
As $f_i^{eq}$ depends on both $\rho$ and $\mvel$ \eqref{eq:feq}, it may be the case that $|f_i^{eq}(\pos, t) - f_i^{eq}(\pos + \pvel_i \Delta t, t + \Delta t)| >> 0$, i.e. there may be a large difference in the quasiequilibrium for the $i^{th}$ discrete velocity at $(\pos, t)$ and $(\pos + \pvel_i \Delta t, t + \Delta t)$.
In this case, if $f_i$ is ``near'' to quasiequilibrium at $\pos$ it will be ``far'' from quasiequilibrium after the streaming step when it moves to the node at $\pos + \pvel_i \Delta t$.
Overrelaxation will occur if $\nu \rightarrow 0$ because consquently $\omega \rightarrow 2$ and \eqref{eq:bgk} results in $f_i(\pos + \pvel_i \Delta t,t + \Delta t)$ still being ``far'' from $f^{eq}_i(\pos + \pvel_i \Delta t,t + \Delta t)$ but on the ``other side''.
Overrelaxation in subsequent time steps (along the streaming trajectory of $f_i$) could result in nonphysical oscillations.
Considering the effect oscillations of particle distributions will have on macroscopic variables and, consequently, local quasiequilibriums, positive feedback loops can occur causing the system to diverge or ``pollute'' the system enough to make the results highly nonphysical~\cite{gorban2014enhancement}.

@@

Due to the instabilities associated with the collision frequency being too high (e.g. approaching 2) or too low (e.g. approaching 0), a natural, albiet nonphysical, approach to using the BGK collision operator for non-Newtonian flow is to simply put bounds on the values in which the collision frequency may attain (such as in~\citet{svec2011flow,svec2012free,gabbanelli2005lattice}).
This simple methodology for increasing stability will henceforth be referred to bounded-relaxation time BGK, or BGK-BRT.

@@

\noindent where $\overline{\Delta S}$ and $\sigma_{\Delta S}$ are the mean and standard deviation of $\Delta S$, respectively--both are calculated using values over the domain for the current timestep--and $n_\sigma$ is the number of standard deviations greater than $\overline{\Delta S}$ that $\Delta S$ must be before dissipation is added.
The number of standard deviations, $n_\sigma$, is chosen \textit{a priori}.
The criteria described in \eqref{eq:dissipation-criteria} has the advantage that one does not need to determine \textit{a priori} what constitutes ``far'' from quasiequilibrium, but instead considers the maximum percentage of sites one would want artificial dissipation to be added to.
A disadvantage of \eqref{eq:dissipation-criteria} is that if $\overline{\Delta S}$ and $n_\sigma$ are both small then it is possible for dissipation to be added when $\Delta S(\pos, t)$ is low and artifical dissipation is unnecessary.
By requiring that both \eqref{eq:ds-threshold} and \eqref{eq:dissipation-criteria} be met before artifical dissipation is added, there is the potential for \eqref{eq:ds-threshold} and \eqref{eq:dissipation-criteria} to compensate for the each other's disadvantage.

@@

No slip, or zero velocity, which is commonly imposed at walls in a domain, is accomplished by simulating the particle distributions as "bouncing back" at the walls in the opposite direction from which they stream.
For example, for a particle distribution streaming in the direction of a south wall, $f_2 = f_4$, $f_5 = f_7$, and $f_6 = f_8$.
For velocity or pressure boundary conditions, the method proposed by~\citet{zou1997pressure} can be used.
The particle distributions that are missing after the streaming step are solved for by assuming a bounceback of the nonequilibrium distribution in the direction normal to the boundary; e.g., for a south inlet or outlet, $f_2 - f_2^{eq} = f_4 - f_4^{eq}$.

@@

In the present work, material behavior is simulated through the collision frequency, $\omega$, when using the BGK collision operator (\Fref{sec:bgk}), or the $s_7$ and $s_8$ relaxation rates when using the MRT collision operator (\Fref{sec:mrt}).
Both the collision frequency and the $s_7$ and $s_8$ relaxation rates are dependent on the apparent visocity, $\mu_{app}$, through \eqref{eq:colfreq-non-newtonian} and \eqref{eq:mrt-s7}, respectively.
The apparent viscosity, in the context of non-Newtonian flow, is a function of the local strain-rate, $\dot{\gamma}$.
The strain-rate is given by the second invariant of the strain-rate tensor, $\srtensor$, i.e.:
\begin{equation} \label{eq:second-invariant}
\dot{\gamma} = \sqrt{2 \sum_{\alpha,\beta=1}^{2} \srtensor \srtensor}.
\end{equation}

@@

Unless otherwise stated, all of the Poiseuille flow simulations were computed on a $32 \times 64$ lattice for 25000 timesteps.
The lattice size was chosen to be sufficiently fine for accuracy and the number of time steps were set high enough to ensure a steady-state would be reached.
In reference to computational time, each of the simulations in this section were run on a single core of an Intel I7-860 Quad-Core 2.80GHz processor.
For the MRT relaxation matrix the free parameters were set to $s_1 = s_2 = s_4 = s_6 = 1.1$, which follows the recommendation of~\cite{lallemand2000theory} for reasons of stability, and has been successfully used to simulate non-Newtonian flow in the past~\cite{fallah2012multiple,grasinger2015simulation}.

@@

As has been reported previously, for the BGK collision operator, using a stress growth exponent of $m = 10^5$ is more accurate with respect to the analytical solution than using a stress growth exponent of $m = 10^8$~\cite{chen2014simulations}.
A larger stress growth exponent results in a more accurate Papastasiou approximation of the true Bingham plastic constitutive model, however, it leads to more nonphysical oscillations that degrade the numerical solution for the BGK collision operator.
Upon inspection of \Fref{tab:poise-bing}, it does not appear that entropic median filtering helped to mitigate errors that occured as a result of using $m = 10^8$ for Bingham plastic Poiseuille flow.
In fact, the median filter resulted in a less accurate solution in all cases other than the case with the lowest yield stress considered (i.e., the smoothest flow field), and rendered the solution altogether useless for the higher yield stress fluids (relative errors of approximately 25--75\%).

@@

In order to better understand why median filtering did not eliminate the nonphysical oscillations in high yield stress fluids, but instead exacerbated the problem, it is necessary to investigate what is happening at the particle distribution scale.
\Fref{fig:feq-vs-f_bgk} compares particle distributions to quasiequilibrium.
More specifically, \Fref{fig:feq-vs-f_bgk} compares $f_i(\pos_j, 25000)$ to $f_i^{eq}(\pos_j, 25000)$ where $\pos_j = \begin{Bmatrix}16 & y_j\end{Bmatrix}^T$ for $j = 1, 2, ..., 64$, and $i = 5, 8$.
It can be seen that in the BGK, $m = 10^8$ solution, that nonphysical oscillations pollute the quasiequilibrium profile as well.
These nonphysical oscillations likely originate at the lattice scale due to the sharp gradient in the macroscopic velocity, $\mvel$, near the walls, and the sharp gradient in $\mvel$ is a result of a sharp gradient in $\mu_{app}$, namely the sharp gradient of the constitutive relationship.
Because the nonphysical oscillations make their way into the quasiequilibrium values, entropic median filtering does not help dampen the oscillations but instead contracts the particle distributions closer to the quasiequilibrium oscillations, which explains why the median filtered results were less accurate for Bingham plastic Poiseuille flow.
In order to ensure that this phenomena was a side effect of all entropic filtering, and not just entropic median filtering with $\theta = 1.0 \times 10^{-6}$ and $n_s = 2.7$, optimization was used to find the value of $\theta$ that minimized error for both median filtering and Ehrenfests' regularization.
The optimization problem was defined as follows:
\begin{align*}
  \min_\theta \hspace{0.2in} & f(\theta)\\
  \text{such that} \hspace{0.2in} & \theta \in \begin{bmatrix}10^{-10}, & 2.0\end{bmatrix}
\end{align*}
\noindent where $f(\theta) \equiv $ the relative $L_2$ error between the LBM approximation using $m = 10^8$ and entropic filtering with a $\Delta S$ threshold of $\theta$ (used in the criteria defined in \eqref{eq:ds-threshold}).
The bounds on $\theta$, namely $\theta \in \begin{bmatrix}10^{-10}, & 2\end{bmatrix}$ represent $\theta$ such that 100\% of lattice sites are filtered for every time step ($\theta = 10^{-10}$) and $\theta$ such that no lattice sites are filtered for any time step ($\theta = 2.0$).
The optimization problem was solved for the cases of using median filtering and Ehrenfests' regularization using both Brent's method and the Golden Section search.
For all four cases, the optimal solution was $\theta = 2.0$ and the cost function, $f(2.0)$, was equal to the relative $L_2$ error for the BGK collision operator, $m = 10^8$ and no entropic filtering.
The results of the optimization show that entropic filtering, at best did not affect the accuracy of LBM, and often adversly affected the accuracy of LBM in approximating Bingham plastic Poiseuille flow, regardless of whether median filtering or Ehrenfests' regularization was used, and regardless of what the $\Delta S$ threshold, $\theta$, was set to.

@@

On average, the LBM model using the MRT collision operator took 2.9 times more computing time than the BGK collision operator with $m = 10^5$.
However, the increased computing time can be justified for the MRT collision operator because the MRT solution did not suffer from the same level of nonphysical oscillations as the BGK collision operator when using the more accurate approximation of the Bingham plastic model, i.e. when $m = 10^8$. 
The MRT collision operator was therefore the most accurate solution for all cases.
\Fref{fig:feq-vs-f_mrt} compares particle distribution, $f_i$, profiles to their respective quasiequilbirium profiles, $f_i^{eq}$ for the MRT model.
In contrast to the BGK with $m = 10^8$, there were no significant nonphysical oscillations in the quasiequilibrium distributions.

@@

% MG: it might be better to normalize the time with respect to the fastest, BGK with no filtering

@@

The lid-driven cavity simulations presented in the section were all simulated on a relatively coarse, $100 \times 100$ lattice, similar to the work presented in ~\cite{brownlee2008nonequilibrium}.
A coarse lattice was chosen in order to highlight concerns with stability and accuracy.
The simulations were run for either 50000 timesteps or until convergence was met.
Convergence was defined by:
\begin{equation} \label{eq:convergence}
\sum_{m=1}^{100} \sum_{i, j} \frac{|u_{i, j}^k - u_{i, j}^{k-m}|}{|u_{i, j}^{k-m}|} < 1.0 \times 10^{-7},
\end{equation}
\noindent where $i$ is the node index in the x-direction, $j$ is the node index in the y-direction, and $k$ is the current timestep.
The lid velocity was prescribed as $u_o = 0.1$.
For the MRT relaxation matrix, the free parameters were set to $s_1 = 1.1$, $s_2 = 1.0$, and $s_4 = s_6 = 1.2$, as these values have been successfully applied to simulating lid-driven cavity flow of non-Newtonian fluids in the past~\cite{chen2014simulations,li2014simulation}.
All coordinate values presented in this section (used to describe the location of the center of vortices) are given normalized with respect to the length of the cavity side (i.e. as $(x / L, y / L)$).
Note that all results presented as ``-'' indicate that the tests were numerically unstable to the extent that the degrees of freedom all over the domain were diverging toward $-\infty$ or $\infty$.

@@

For flow with $Re \ge 5000$, the BGK with $m = 10^8$ and median filtering produced numerically stable results.
However, in general, the BGK with $m = 10^8$ and median filtering was consistently different from literature values. 
The apparent inaccuracy with regards to BGK with $m = 10^8$ and median filtering is probably due to how the numerical stability is enhanced through artificial, nonphysical dissipation.
In contrast to the nonphysical nature in which median filtering enhances stability, the stability enhancement used by MRT does not directly affect the macroscopic hydrodynamics of interest, so it again provides a stable, and in most of the cases examined herein, the most accurate solution. % MG: verify this sentence is correct
The apparent superiority in terms of stability and accuracy of the MRT collision operator still comes at a price though.
The MRT collision operator was, in general, 5-10 times slower than any of the BGK collision schemes.
The extreme increase in computational expense is probably not due to the collision process itself (i.e. \eqref{eq:mrt-colop}) but instead due to calculating the strain-rate \eqref{eq:srtensor-mrt} for each iteration of the solution for the apparent viscosity, $\mu_{app}$.

@@

For the shear-thinning results, $n = 0.5$, at the lowest Reynold's number considered, $Re = 100$, all of the collision schemes produced results that agreed well with the literature value.
However, the solutions became unstable at higher Reynold's numbers and entropic filtering was necessary to produce a stable soltuion when $Re \ge 1000$ for the BGK collision operator, and when $Re \ge 5000$ for the MRT collision operator.
Although there still appears to be a tradeoff between computational efficiency and accuracy in regards to the BGK with median filtering and MRT with median filtering, it does not seem to be nearly as significant as when simulating Bingham plastic lid-driven flow (e.g. neither collision scheme consistently agrees with the literature values and the MRT with median filtering is only 2-3 times slower than the BGK with median filtering).
An example of this tradeoff is the shear-thinning results for $Re = 5000$: the location of the main vortex for the BGK with median filtering differed from the literature by more than 10\% in both the $x$ and $y$ directions, but the MRT with median filtering has approximately twice the computing cost.

@@

In regards to future work, \cite{conrad2015accuracy} showed that non-Newtonian Lattice Boltzmann simulations are not second order accurate for every simulation Mach number, and showed that a greater accuracy may be achieved by properly tuning the simulation Mach number.
\cite{conrad2015accuracy} predicted optimal simulation Mach numbers for various shear-thinning and shear-thickening fluids. 
Possbile future work would include extending the numerical investigation to consider the effect of the simulation Mach number on the accuracy, stability, and efficiency of the various collision operators and stability enhancements in simulating non-Newtonian flow.

>>>>

There are a number of fluids in science and engineering applications that can be classified as non-Newtonian, such as pastes, slurries, molten plastics, polymer solutions, dyes, varnishes, suspensions, and some biomedical liquids such as blood all behave in a non-Newtonian manner~\cite{bohme1987non}.
Of all of the different non-Newtonian behaviors that exist, there are two models under which much of the behaviors may be idealized: yield stress fluids and power-law fluids.
Yield stress fluids, also known as Bingham plastics, do not flow until a threshold value of stress, referred to as its yield stress, is exceeded.
Yield stress flow has been found to be relevant in many applications, such as flow of pastes, paints, muds, molten plastics and metals, and in some cases, blood~\cite{wang2011lattice}.

@@

Due to its simplicity and computational efficiency, the most common collision operator is the Bhatnagar--Gross--Krook (BGK) operator.
BGK consists of a single relaxation time and is a linear relaxation of particle distributions toward equilibrium.
The BGK collision operator for the $i^{th}$ discrete velocity is expressed as:

@@

Despite the utility of the BGK collision operator, it does have a few drawbacks.
For example, in low viscosity fluids the BGK operator results in an overrelaxation of particle distributions toward quasiequilibrium.
It is well known that when large nonequilibrium distributions exist in the LBM approximation that overrelaxation can result in nonphysical oscillations that are slow to decay~\cite{brownlee2007stability,dellar2003incompressible}.
To illustrate this, consider a flow in which there is a sharp spatial gradient in either $\rho$ or $\mvel$ at $\pos$.
As $f_i^{eq}$ depends on both $\rho$ and $\mvel$ \eqref{eq:feq}, it may be the case that $|f_i^{eq}(\pos, t) - f_i^{eq}(\pos + \pvel_i \Delta t, t + \Delta t)| >> 0$, i.e. there may be a large difference in the quasiequilibrium for the $i^{th}$ discrete velocity at $(\pos, t)$ and $(\pos + \pvel_i \Delta t, t + \Delta t)$.
In this case, if $f_i$ is ``near'' to quasiequilibrium at $\pos$ it will be ``far'' from quasiequilibrium after the streaming step when it moves to the node at $\pos + \pvel_i \Delta t$.
Overrelaxation will occur if $\nu \rightarrow 0$ because consequently $\omega \rightarrow 2$ and \eqref{eq:bgk} results in $f_i(\pos + \pvel_i \Delta t,t + \Delta t)$ still being ``far'' from $f^{eq}_i(\pos + \pvel_i \Delta t,t + \Delta t)$ but on the ``other side''.
Overrelaxation in subsequent time steps (along the streaming trajectory of $f_i$) could result in nonphysical oscillations.
Considering the effect oscillations of particle distributions will have on macroscopic variables and, consequently, local quasiequilibriums, positive feedback loops can occur causing the system to diverge or ``pollute'' the system enough to make the results highly nonphysical~\cite{gorban2014enhancement}.

@@

Due to the instabilities associated with the collision frequency being too high (e.g. approaching 2) or too low (e.g. approaching 0), a natural, albeit nonphysical, approach to using the BGK collision operator for non-Newtonian flow is to simply put bounds on the values in which the collision frequency may attain (such as in~\citet{svec2011flow,svec2012free,gabbanelli2005lattice}).
This simple methodology for increasing stability will henceforth be referred to bounded-relaxation time BGK, or BGK-BRT.

@@

\noindent where $\overline{\Delta S}$ and $\sigma_{\Delta S}$ are the mean and standard deviation of $\Delta S$, respectively--both are calculated using values over the domain for the current time step--and $n_\sigma$ is the number of standard deviations greater than $\overline{\Delta S}$ that $\Delta S$ must be before dissipation is added.
The number of standard deviations, $n_\sigma$, is chosen \textit{a priori}.
The criteria described in \eqref{eq:dissipation-criteria} has the advantage that one does not need to determine \textit{a priori} what constitutes ``far'' from quasiequilibrium, but instead considers the maximum percentage of sites one would want artificial dissipation to be added to.
A disadvantage of \eqref{eq:dissipation-criteria} is that if $\overline{\Delta S}$ and $n_\sigma$ are both small then it is possible for dissipation to be added when $\Delta S(\pos, t)$ is low and artificial dissipation is unnecessary.
By requiring that both \eqref{eq:ds-threshold} and \eqref{eq:dissipation-criteria} be met before artificial dissipation is added, there is the potential for \eqref{eq:ds-threshold} and \eqref{eq:dissipation-criteria} to compensate for the each other's disadvantage.

@@

No slip, or zero velocity, which is commonly imposed at walls in a domain, is accomplished by simulating the particle distributions as "bouncing back" at the walls in the opposite direction from which they stream.
For example, for a particle distribution streaming in the direction of a south wall, $f_2 = f_4$, $f_5 = f_7$, and $f_6 = f_8$.
For velocity or pressure boundary conditions, the method proposed by~\citet{zou1997pressure} can be used.
The particle distributions that are missing after the streaming step are solved for by assuming a bounce back of the nonequilibrium distribution in the direction normal to the boundary; e.g., for a south inlet or outlet, $f_2 - f_2^{eq} = f_4 - f_4^{eq}$.

@@

In the present work, material behavior is simulated through the collision frequency, $\omega$, when using the BGK collision operator (\Fref{sec:bgk}), or the $s_7$ and $s_8$ relaxation rates when using the MRT collision operator (\Fref{sec:mrt}).
Both the collision frequency and the $s_7$ and $s_8$ relaxation rates are dependent on the apparent viscosity, $\mu_{app}$, through \eqref{eq:colfreq-non-newtonian} and \eqref{eq:mrt-s7}, respectively.
The apparent viscosity, in the context of non-Newtonian flow, is a function of the local strain-rate, $\dot{\gamma}$.
The strain-rate is given by the second invariant of the strain-rate tensor, $\srtensor$, i.e.:
\begin{equation} \label{eq:second-invariant}
\dot{\gamma} = \sqrt{2 \sum_{\alpha,\beta=1}^{2} \srtensor \srtensor}.
\end{equation}

@@

Unless otherwise stated, all of the Poiseuille flow simulations were computed on a $32 \times 64$ lattice for 25000 time steps.
The lattice size was chosen to be sufficiently fine for accuracy and the number of time steps were set high enough to ensure a steady-state would be reached.
In reference to computational time, each of the simulations in this section were run on a single core of an Intel I7-860 Quad-Core 2.80GHz processor.
For the MRT relaxation matrix the free parameters were set to $s_1 = s_2 = s_4 = s_6 = 1.1$, which follows the recommendation of~\cite{lallemand2000theory} for reasons of stability, and has been successfully used to simulate non-Newtonian flow in the past~\cite{fallah2012multiple,grasinger2015simulation}.

@@

As has been reported previously, for the BGK collision operator, using a stress growth exponent of $m = 10^5$ is more accurate with respect to the analytical solution than using a stress growth exponent of $m = 10^8$~\cite{chen2014simulations}.
A larger stress growth exponent results in a more accurate Papastasiou approximation of the true Bingham plastic constitutive model, however, it leads to more nonphysical oscillations that degrade the numerical solution for the BGK collision operator.
Upon inspection of \Fref{tab:poise-bing}, it does not appear that entropic median filtering helped to mitigate errors that occurred as a result of using $m = 10^8$ for Bingham plastic Poiseuille flow.
In fact, the median filter resulted in a less accurate solution in all cases other than the case with the lowest yield stress considered (i.e., the smoothest flow field), and rendered the solution altogether useless for the higher yield stress fluids (relative errors of approximately 25--75\%).

@@

In order to better understand why median filtering did not eliminate the nonphysical oscillations in high yield stress fluids, but instead exacerbated the problem, it is necessary to investigate what is happening at the particle distribution scale.
\Fref{fig:feq-vs-f_bgk} compares particle distributions to quasiequilibrium.
More specifically, \Fref{fig:feq-vs-f_bgk} compares $f_i(\pos_j, 25000)$ to $f_i^{eq}(\pos_j, 25000)$ where $\pos_j = \begin{Bmatrix}16 & y_j\end{Bmatrix}^T$ for $j = 1, 2, ..., 64$, and $i = 5, 8$.
It can be seen that in the BGK, $m = 10^8$ solution, that nonphysical oscillations pollute the quasiequilibrium profile as well.
These nonphysical oscillations likely originate at the lattice scale due to the sharp gradient in the macroscopic velocity, $\mvel$, near the walls, and the sharp gradient in $\mvel$ is a result of a sharp gradient in $\mu_{app}$, namely the sharp gradient of the constitutive relationship.
Because the nonphysical oscillations make their way into the quasiequilibrium values, entropic median filtering does not help dampen the oscillations but instead contracts the particle distributions closer to the quasiequilibrium oscillations, which explains why the median filtered results were less accurate for Bingham plastic Poiseuille flow.
In order to ensure that this phenomena was a side effect of all entropic filtering, and not just entropic median filtering with $\theta = 1.0 \times 10^{-6}$ and $n_s = 2.7$, optimization was used to find the value of $\theta$ that minimized error for both median filtering and Ehrenfests' regularization.
The optimization problem was defined as follows:
\begin{align*}
  \min_\theta \hspace{0.2in} & f(\theta)\\
  \text{such that} \hspace{0.2in} & \theta \in \begin{bmatrix}10^{-10}, & 2.0\end{bmatrix}
\end{align*}
\noindent where $f(\theta) \equiv $ the relative $L_2$ error between the LBM approximation using $m = 10^8$ and entropic filtering with a $\Delta S$ threshold of $\theta$ (used in the criteria defined in \eqref{eq:ds-threshold}).
The bounds on $\theta$, namely $\theta \in \begin{bmatrix}10^{-10}, & 2\end{bmatrix}$ represent $\theta$ such that 100\% of lattice sites are filtered for every time step ($\theta = 10^{-10}$) and $\theta$ such that no lattice sites are filtered for any time step ($\theta = 2.0$).
The optimization problem was solved for the cases of using median filtering and Ehrenfests' regularization using both Brent's method and the Golden Section search.
For all four cases, the optimal solution was $\theta = 2.0$ and the cost function, $f(2.0)$, was equal to the relative $L_2$ error for the BGK collision operator, $m = 10^8$ and no entropic filtering.
The results of the optimization show that entropic filtering, at best did not affect the accuracy of LBM, and often adversely affected the accuracy of LBM in approximating Bingham plastic Poiseuille flow, regardless of whether median filtering or Ehrenfests' regularization was used, and regardless of what the $\Delta S$ threshold, $\theta$, was set to.

@@

On average, the LBM model using the MRT collision operator took 2.9 times more computing time than the BGK collision operator with $m = 10^5$.
However, the increased computing time can be justified for the MRT collision operator because the MRT solution did not suffer from the same level of nonphysical oscillations as the BGK collision operator when using the more accurate approximation of the Bingham plastic model, i.e. when $m = 10^8$. 
The MRT collision operator was therefore the most accurate solution for all cases.
\Fref{fig:feq-vs-f_mrt} compares particle distribution, $f_i$, profiles to their respective quasiequilibrium profiles, $f_i^{eq}$ for the MRT model.
In contrast to the BGK with $m = 10^8$, there were no significant nonphysical oscillations in the quasiequilibrium distributions.

@@

@@

The lid-driven cavity simulations presented in the section were all simulated on a relatively coarse, $100 \times 100$ lattice, similar to the work presented in ~\cite{brownlee2008nonequilibrium}.
A coarse lattice was chosen in order to highlight concerns with stability and accuracy.
The simulations were run for either 50000 time steps or until convergence was met.
Convergence was defined by:
\begin{equation} \label{eq:convergence}
\sum_{m=1}^{100} \sum_{i, j} \frac{|u_{i, j}^k - u_{i, j}^{k-m}|}{|u_{i, j}^{k-m}|} < 1.0 \times 10^{-7},
\end{equation}
\noindent where $i$ is the node index in the x-direction, $j$ is the node index in the y-direction, and $k$ is the current time step.
The lid velocity was prescribed as $u_o = 0.1$.
For the MRT relaxation matrix, the free parameters were set to $s_1 = 1.1$, $s_2 = 1.0$, and $s_4 = s_6 = 1.2$, as these values have been successfully applied to simulating lid-driven cavity flow of non-Newtonian fluids in the past~\cite{chen2014simulations,li2014simulation}.
All coordinate values presented in this section (used to describe the location of the center of vortices) are given normalized with respect to the length of the cavity side (i.e. as $(x / L, y / L)$).
Note that all results presented as ``-'' indicate that the tests were numerically unstable to the extent that the degrees of freedom all over the domain were diverging toward $-\infty$ or $\infty$.

@@

For flow with $Re \ge 5000$, the BGK with $m = 10^8$ and median filtering produced numerically stable results.
However, in general, the BGK with $m = 10^8$ and median filtering was consistently different from literature values. 
The apparent inaccuracy with regards to BGK with $m = 10^8$ and median filtering is probably due to how the numerical stability is enhanced through artificial, nonphysical dissipation.
In contrast to the nonphysical nature in which median filtering enhances stability, the stability enhancement used by MRT does not directly affect the macroscopic hydrodynamics of interest, so it again provides a stable, and in most of the cases examined herein, the most accurate solution.
The apparent superiority in terms of stability and accuracy of the MRT collision operator still comes at a price though.
The MRT collision operator was, in general, 5-10 times slower than any of the BGK collision schemes.
The extreme increase in computational expense is probably not due to the collision process itself (i.e. \eqref{eq:mrt-colop}) but instead due to calculating the strain-rate \eqref{eq:srtensor-mrt} for each iteration of the solution for the apparent viscosity, $\mu_{app}$.

@@

For the shear-thinning results, $n = 0.5$, at the lowest Reynold's number considered, $Re = 100$, all of the collision schemes produced results that agreed well with the literature value.
However, the solutions became unstable at higher Reynold's numbers and entropic filtering was necessary to produce a stable solution when $Re \ge 1000$ for the BGK collision operator, and when $Re \ge 5000$ for the MRT collision operator.
Although there still appears to be a tradeoff between computational efficiency and accuracy in regards to the BGK with median filtering and MRT with median filtering, it does not seem to be nearly as significant as when simulating Bingham plastic lid-driven flow (e.g. neither collision scheme consistently agrees with the literature values and the MRT with median filtering is only 2-3 times slower than the BGK with median filtering).
An example of this tradeoff is the shear-thinning results for $Re = 5000$: the location of the main vortex for the BGK with median filtering differed from the literature by more than 10\% in both the $x$ and $y$ directions, but the MRT with median filtering has approximately twice the computing cost.

@@

In regards to future work, \cite{conrad2015accuracy} showed that non-Newtonian Lattice Boltzmann simulations are not second order accurate for every simulation Mach number, and showed that a greater accuracy may be achieved by properly tuning the simulation Mach number.
\cite{conrad2015accuracy} predicted optimal simulation Mach numbers for various shear-thinning and shear-thickening fluids. 
Possible future work would include extending the numerical investigation to consider the effect of the simulation Mach number on the accuracy, stability, and efficiency of the various collision operators and stability enhancements in simulating non-Newtonian flow.
